Minimax optimal Laplacian smoothing

### 2020-10-06 notes/manuscripts/AISTATS/minimax_optimal_laplacian_smoothing2.tex
Fixed a few typos, and shortened one wandering paragraph.

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green, Sivaraman Balakrishnan, and Ryan Tibshirani}
\date{\today}
\maketitle

@@

Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}

@@

Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of \emph{fixed graph} $G$. and bound the statistical error as a function of $\theta$ and $G$. (In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise. As such, both our analysis and conclusions are naturally quite different.

Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit. They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
\begin{equation}
\label{eqn:laplace_operator}
\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
\end{equation}
as $n \to \infty$ and $r \to 0$. This is meant in various senses:
\begin{itemize}
	\item Pointwise convergence, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}. 
	\item Spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
		$$
		\Delta_Pg = \lambda g,\\
		\restr{g}{\partial \Xset} = 0
		$$} 
	\item Convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. 
\end{itemize}
While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.

@@

\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
\begin{itemize}
	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$. We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, and conclude in Section~\ref{sec:discussion} with some potential methodological and theoretical extensions.
\end{itemize}

@@

@@

Before we get to our main results, we briefly review minimax optimal estimation and testing rates over Sobolev classes. We also single out a notable method which achieves these rates: \emph{smoothing splines}. As we will see, Laplacian smoothing can be seen as a graph-based---i.e. discrete---counterpart to the ``continuous-time'' approach of smoothing splines.

@@

When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory. However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting. More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.

For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$. However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator: it has a nice closed-form solution amenable to a bias-variance decomposition. We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.

\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
To better motivate Laplacian smoothing as a tool for nonparametric regression, we review its connection to one of the most classical methods of nonparametric regression: smoothing splines.

Smoothing splines are specially designed for regression over Sobolev spaces: they are penalized least squares estimators where the penalty is itself the Sobolev semi-norm. Mathematically, the smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
\end{equation}
In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.

By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wh{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$. Suppose for the moment that $P$ is the uniform distribution over $\Xset$. Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
\begin{equation}
\label{eqn:seminorm_convergence}
\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
\end{equation}

@@

Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
\begin{equation}
\label{eqn:sobolev_space_estimation_minimax_rate}
\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
\end{equation}
In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
\end{equation} 
where $\mc{F}$ is some class of functions. The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$. Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
\begin{equation}
\label{eqn:sobolev_space_testing_critical_radius}
\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
\end{equation}
see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in goodness-of-fit testing. When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.

The estimator $\wh{f}_{\SM}$ and the level-$\alpha$ test $\phi_{\SM} := \1\{T_{\SM} \geq {t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g. \cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result). When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: as illustrated in e.g. \cite{green93}, the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}_{\SM}$ will be inconsistent. By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and \cite{birge1993} show that the resulting Lipschitz smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem. We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Lipschitz smoothing spline has risk provably larger than optimal \cite{birge1993}. Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$, but interestingly the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.

Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}. Encouragingly, the smoothing spline estimator $\wt{f}$ and test $\wt{\phi}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.

\paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former. Unfortunately, it is not so easy to mathematically tie $\wh{f}$ to $\wt{f}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$. Instead, we will take a different approach and analyze $\wh{f}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.

@@

To understand the role graph Sobolev balls play in our analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting. Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$. The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f|_{H^1(\Xset)}^2 = \langle f, \Delta_P f \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  only for functions $f \in C^2(\Xset)$ which satisfy boundary conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e. Weyl's Law). In this section we will show that both these statements have analogues in the graph setting.
\begin{itemize}
	\item For functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm_vague}
	f^T \Lap_{n,r} f \leq C M^2(n^{2}r^{(d + 2)}),
	\end{equation}
	\item For a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
	\begin{equation}
	\label{eqn:neighborhood_graph_eigenvalue_vague}
	ck^{2/d} \leq \frac{1}{nr^{d + 2}} \lambda_k(G_{n,r})  \leq Ck^{2/d}.
	\end{equation}
\end{itemize}
Both statements will hold with high probability, and when we assume $\Xset$ is a manifold of intrinsic dimension $m$, they will hold with $d$ replaced by $m$.  These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.

>>>>

\begin{document}
\title{Minimax optimal Laplacian smoothing}
\author{Alden Green, Sivaraman Balakrishnan, and Ryan J. Tibshirani}
\date{\today}
\maketitle

@@

Laplacian smoothing has many advantages compared to other nonparametric regression methods. For instance:
\begin{itemize}
	\item It is fast, easy, and stable to compute. Letting ${\bf Y} = (Y_1,\ldots,Y_n) \in \Reals^n$, it is not hard to see that the minimizer $\wt{f}_{\LS} = (\rho \Lap_{n,r} + I)^{-1}{\bf Y}$ can be computed by solving a linear equation; in practice, one typically chooses $r$ to be small, with the result being that the matrix $\Lap_{n,r}$ is sparse. There exist known fast solvers of exactly this system.
	\item Is is generalizable to non-standard data---for instance, text or images, or really any data modality on which it is possible to define a kernel.
	\item It is easily adaptable to the semi-supervised learning setting. 
\end{itemize}
\textcolor{red}{(TODO)}: Perhaps additional motivation is needed?

@@

Works in the first category view the responses $Y = \theta + \varepsilon$ as noisy samples of a regression function $\theta \in \Reals^n$ defined on the vertices of a \emph{fixed graph} $G$. and bound the statistical error as a function of $\theta$ and $G$. (In the estimation literature, see for instance \citep{wang2016, sadhanala16,sadhanala17,kirichenko2017,kirichenko2018}; in the testing literature, see \citep{sharpnack2013,sharpnack2013b,sharpnack2015}.) Note that in our setting, the graph $G_{n,r}$ and the evaluations $(f_0(X_1),\ldots,f_0(X_n))$ are random rather than fixed, and our assumptions are placed only on the underlying objects $P$ and $f_0$ from which these random object arise. As such, both our analysis and conclusions are naturally quite different.

Works in the second category prove the consistency of neighborhood graph Laplacian smoothing and other neighborhood graph-based methods in the large-sample limit. They do so by establishing the convergence of the graph Laplacian $\Lap_{{n,r}}$ to the continuum Laplace operator $\Delta_P: C^2(\Xset) \to \Leb^2(\Xset)$,
\begin{equation}
\label{eqn:laplace_operator}
\Delta_Pg := -\frac{1}{p} \dive(p^2 \nabla g)
\end{equation}
as $n \to \infty$ and $r \to 0$. This is meant in (at least) three senses: (1) \emph{pointwise convergence}, meaning that $\Lap_{n,r}f \to \Delta_Pf$ in an appropriate norm \citep{belkin03,belkin05,lafon04,hein05,singer06,gine06}; (2) spectral convergence, meaning the eigenvalues $\lambda_k(\Lap_{n,r}) \to \lambda_k(\Delta_P)$, where the latter are the Dirichlet eigenvalues of $\Delta_P$ \citep{belkin07,burago2014,trillos2018,trillos2019,calder2019}.\footnote{\label{footnote:dirichlet_eigenvalues} Formally, Dirichlet eigenvalues are solutions to the problem
	$$
	\Delta_Pg = \lambda g,\\
	\restr{g}{\partial \Xset} = 0
	$$}, and (3) convergence of norms, meaning $f^T \Lap_{n,r} f \to \langle f,\Delta_Pf \rangle_{\Leb^2(P)}$ \citep{bousquet03,hein06,zhou11}. While these results imply \emph{consistency} of many different graph Laplacian based procedures, the statistical \emph{optimality} of Laplacian smoothing over neighborhood graphs is still not well understood.

@@

\paragraph{Organization.} We now describe the structure of the rest of this paper, which builds to the aforementioned minimax results.
\begin{itemize}
	\item In Section~\ref{sec:minimax_optimal_regression_sobolev_spaces}, for completeness we review some standard results regarding minimax nonparametric regression over Sobolev spaces.
	\item In Section~\ref{sec:graph_sobolev_classes}, we introduce the graph Sobolev class $H^1(G_{n,r})$, which we view as a graph-based analog to the continuum Sobolev class $H^1(\Xset)$. We show that key properties of $H^1(\Xset)$ used to derive minimax rates are shared by $H^1(G_{n,r})$.
	\item In Section~\ref{sec:minimax_optimal_laplacian_smoothing}, we leverage the results of Section~\ref{sec:graph_sobolev_classes} to derive tight upper bounds on the error of Laplacian smoothing methods.
	\item In Section~\ref{sec:simulations} we provide some experiments supporting our theory, before concluding in Section~\ref{sec:discussion}.
\end{itemize}

@@

\textcolor{red}{(TODO)}: Complete this.

@@

Before we get to our main results, we review minimax optimal estimation and testing rates over Sobolev classes. We also single out a notable method which achieves these rates: \emph{smoothing splines}. As we will see, Laplacian smoothing can be seen as a graph-based---i.e. discrete---counterpart to the ``continuous-time'' approach of smoothing splines.

@@

% AJG commented this out.
% When $d = 1$, the embedding of $H^1(\Xset)$ into $C^{0,q}(\Xset)$ facilitates a general analysis of (penalized) least estimators over $H^1(\Xset)$, using arguments from empirical process theory. However when $d \geq 2$, such arguments fall apart: for example, good bounds on the metric entropy of $H^1(\Xset)$ are not known in this setting. More troubling still, least squares estimators over $H^1(\Xset)$ are not even well-posed when $d \geq 2$.

% For these reasons, analysis of least squares estimators for the nonparametric regression problem over Sobolev spaces typically assume $d = 1$---or more generally that $f_0 \in H^s(\Xset)$ for some $2s > d$. However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator---it has a nice closed-form solution amenable to a bias-variance decomposition. We therefore do not need to rely on empirical process theory, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.

As we will see, the Sobolev embedding theorem has rather grave consequences for least-squares estimators over $H^1(\Xset)$ when $d > 1$, rendering such estimators ill-posed. However, the Laplacian smoother $\wh{f}_{\LS}$ is a special least-squares estimator. It has a nice closed-form solution amenable to a bias-variance decomposition, and so we will be able to analyze $\wh{f}_{\LS}$ and $T_{\LS}$ beyond the univariate setting.

\paragraph{Connecting Laplacian smoothing to smoothing splines.} 
To better motivate Laplacian smoothing as a tool for nonparametric regression, we connect it to one of the most classical methods of nonparametric regression: smoothing splines.

Smoothing splines are specially designed for regression over Sobolev spaces, as they are penalized least squares estimators where the penalty is itself a Sobolev semi-norm. Mathematically, the (first-order) smoothing spline estimator $\wh{f}_{\SM}$ is defined as the solution to the following optimization problem,
\begin{equation}
\label{eqn:smoothing_spline}
\wt{f}_{\SM} = \argmin_{f} \biggl\{\sum_{i = 1}^{n} \bigl(Y_i - f(X_i)\bigr)^2 + \rho \cdot \seminorm{f}_{H^{1}(\Xset)}^2\biggr\}.
\end{equation}
In~\eqref{eqn:smoothing_spline}, the minimum can either be taken over all $f \in H^1(\Xset)$ or over all $f \in C^1(\Xset)$; we shall refer to the former as smoothing splines, and the latter as \emph{Lipschitz} smoothing splines.

By viewing $\Delta_P$ as the continuum limit of~$\Lap_{n,r}$, one can tie the penalty used by the smoothing spline estimator $\wt{f}_{\SM}$ to that used by the Laplacian smoothing estimator $\wh{f}_{\LS}$. Suppose for the moment that $P$ is the uniform distribution over $\Xset$. Using the definition~\eqref{eqn:laplace_operator}, integrating by parts, and invoking the aforementioned convergence properties of $\Lap_{n,r}$, we have that for any $f \in C^{2}(\Xset)$ which satisfies Dirichlet boundary conditions (see footnote~\ref{footnote:dirichlet_eigenvalues}),
\begin{equation}
\label{eqn:seminorm_convergence}
\abs{f}_{H^1(\Xset)}^2 = \int_{\Xset} \bigl(\Delta_P f\bigr)(x) \cdot f(x) \,dx = \lim_{n \to \infty} \frac{1}{n^{2}r^{(d + 2)}}f^T \Lap_{n,r} f.
\end{equation}

@@

Under~\ref{asmp:domain}, the minimax estimation rate over Sobolev balls is (see e.g. \citep{tsybakov2008_book})
\begin{equation}
\label{eqn:sobolev_space_estimation_minimax_rate}
\inf_{\wh{f}} \sup_{f_0 \in H^1(\Xset, M)} \Ebb\Bigl[\norm{\wh{f} - f_0}_{L^2(\Xset)}^2\Bigr] \asymp M^{2d/(2 + d)}n^{-2/(2 + d)}
\end{equation}
In the goodness-of-fit testing problem, we ask for a test function which can distinguish between the null and alternative hypotheses
\begin{equation}
\mathbf{H}_0: f_0 = 0, ~~\textrm{versus}~~ \mathbf{H}_a: f_0 \in \mc{F} \setminus \{0\}
\end{equation} 
where $\mc{F}$ is some class of functions. The minimax critical radius is the smallest value of $\epsilon$ such that some level-${\alpha}$ test $\phi$ has power at least $1 - \alpha$ over all $\mc{F}_{\epsilon} := \mc{F} \cap \{f: \|f\|_{\Leb^2(\Xset)} \geq \epsilon\}$. Testing whether a regression function $f_0$ is equal to $0$ is an easier problem than estimating $f_0$, and so the minimax testing critical radius over $H^1(\Xset,M)$ is much smaller than the minimax estimation rate:
\begin{equation}
\label{eqn:sobolev_space_testing_critical_radius}
\epsilon^2\Bigl(H^1(\Xset,M)\Bigr):= \inf\Biggl\{\epsilon^2: \inf_{\phi} \biggl[\Ebb_0[\phi] +  \sup_{f_0 \in H_{\epsilon}^1(\Xset,M)} \Ebb_{f_0}[1 - \phi]\biggr] \leq 2\alpha \Biggr\} \asymp M^{2d/(4 + d)}n^{-4/(4 + d)}~~\textrm{for $1 \leq d < 4$;}
\end{equation}
see \cite{ingster09} for a proof of this statement, and \cite{ariascastro2018} for a more extended discussion of the minimax paradigm in nonparametric testing. When $d \geq 4$ we no longer have a continuous embedding of $H^1(\Xset)$ into $\Leb^4(\Xset)$; the functions in $H^1(\Xset)$ are very irregular, and the minimax rates in this regime are unknown.

The estimator $\wt{f}_{\SM}$ and the level-$\alpha$ test $\wt{\phi}_{\SM} := \1\{T_{\SM} \geq \wt{t}_{\alpha}\}$ are known to be minimax rate-optimal---meaning they achieve the rates in~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius}, respectively---when $d = 1$ (see e.g. \cite{vandergeer2000} for the estimation result, and \cite{liu2019} for the testing result). When $d > 1$, as a consequence of the Sobolev Embedding Theorem\footnote{Or to be more precise, as a consequence of the \emph{tightness} of the Sobolev Embedding Theorem.} the variational problem~\eqref{eqn:smoothing_spline} is not well-posed: as illustrated in e.g. \cite{green93}, the criterion is minimized by interpolating the responses $\mathbf{Y}$, and the resulting estimator $\wt{f}_{\SM}$ will be inconsistent. By replacing the domain of minimization $H^1(\Xset)$ by $C^1(\Xset)$, we can force the solution to~\eqref{eqn:smoothing_spline} to be less wild, and \cite{birge1993} show that the resulting Lipschitz smoothing spline is not merely consistent but in fact minimax optimal when $d = 2$.\footnote{At least in the estimation problem. We could not find either a positive or negative result for the testing problem.} However when $d > 2$, the Holder space $C^1(\Xset)$ is itself too large, and even the Lipschitz smoothing spline has risk provably larger than optimal \cite{birge1993}. Previewing things a bit, we will see in Section~\ref{sec:minimax_optimal_laplacian_smoothing} that a similar phenomenon occurs for the Laplacian smoother $\wh{f}$, but interestingly the breakdown point will come when $d > 4$, rather than $d > 1$ or $d > 2$.

Finally, it is well known that if $\Xset$ satisfies the manifold hypothesis, meaning informally that it has intrinsic dimension $m < d$ (see~\ref{asmp:domain_manifold} in Section~\ref{sec:graph_sobolev_classes} for a formal statement), both~\eqref{eqn:sobolev_space_estimation_minimax_rate} and~\eqref{eqn:sobolev_space_testing_critical_radius} continue to hold, but with the ambient dimension $d$ replaced by the intrinsic dimension $m$ \citep{bickel2007,ingster2000}. Encouragingly, the smoothing spline estimator $\wt{f}_{\SM}$ and test $\wt{\phi}_{\SM}$ achieve these faster rates, with one caveat: in order for smoothing splines to properly leverage the manifold hypothesis, the domain $\Xset$ must be known.

\paragraph{Consequences?} Intuitively, the connections between Laplacian smoothing and smoothing splines give us reason to hope that the strong statistical properties of the latter might be shared by the former. Unfortunately, it is not so easy to mathematically tie $\wh{f}_{\LS}$ to $\wt{f}_{\SM}$: for example, note that the convergence in~\eqref{eqn:seminorm_convergence} assumes $f \in C^2(\Xset)$ whereas the domain of minimization in~\eqref{eqn:smoothing_spline} is either $H^1(\Xset)$ or $C^1(\Xset)$. Instead, we will take a different approach and analyze $\wh{f}_{\LS}$ by leveraging properties of the graph Sobolev class $H^1(G_{n,r})$, which we now introduce.

@@

To understand the role graph Sobolev balls play in our analysis of Laplacian smoothing, let us return for a moment to the continuous-time setting. Upper bounds on the error of smoothing splines rest on two crucial facts, both of which involve the continuum Laplace operator $\Delta_P$. The first is the \textit{a priori} assumption that the Sobolev semi-norm $|f_0|_{H^1(\Xset)}^2 = \langle f_0, \Delta_P f_0 \rangle_{\Leb^2(P)} \leq M^2$.\footnote{The equality holding when $P$ is uniform and  for functions $f_0 \in C^2(\Xset)$ which satisfy Dirichlet conditions.} The second is an upper bound on the metric entropy of $H^1(\Xset,M)$, characterized by the eigenvalue decay $\lambda_k(\Delta_P) \asymp k^{2/d}$ (i.e. Weyl's Law). In this section we will show that both these statements have analogues in the graph setting.
\begin{itemize}
	\item For functions $f$ belonging to $H^1(\Xset,M)$, the evaluations $f \in (f(X_1),\ldots,f(X_n)) \in \Reals^n$ are upper bounded,
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm_vague}
	f^T \Lap_{n,r} f \leq C M^2(n^{2}r^{(d + 2)}),
	\end{equation}
	\item For a range $1 \leq k \leq k_{\max}(n)$, the eigenvalues $\lambda_k(G_{n,r})$ satisfy
	\begin{equation}
	\label{eqn:neighborhood_graph_eigenvalue_vague}
	ck^{2/d} \leq \frac{1}{nr^{d + 2}} \lambda_k(G_{n,r})  \leq Ck^{2/d}.
	\end{equation}
\end{itemize}
Both statements will hold with high probability, and when we assume $\Xset$ is a manifold of intrinsic dimension $m$, they will hold with $d$ replaced by $m$.  These results elucidate why Laplacian smoothing ``works'' as well as its continuous-time counterpart ---in the sense of being statistically rate-optimal over (continuum) Sobolev classes.

\paragraph{Graph Sobolev semi-norm.}
We begin by upper bounding $f^T \Lap_{n,r} f$ under the assumption $f_0 \in H^1(\Xset,M)$. We will assume the kernel $K$ satisfies condition~\ref{asmp:kernel}.
\begin{enumerate}[label=(K\arabic*)]
	\item
	\label{asmp:kernel}
	$K:[0,\infty) \to [0,\infty)$ is a non-increasing function supported on $[0,1]$, its restriction to $[0,1]$ is Lipschitz, and $K(1) > 0$. Additionally, it is normalized so that
	\begin{equation*}
	\int_{\Reals^d} K\bigl(\norm{z}\bigr) \,dz = 1.
	\end{equation*}
	Let $\sigma_K := \frac{1}{d} \int_{\Rd} \|x\|^2 K(\|x\|) \,dx$.
\end{enumerate}
This is a relatively mild assumption: the choice of kernel is under the user's control, and moreover~\ref{asmp:kernel} covers many (though certainly not all) common kernel choices. In Lemma~\ref{lem:graph_sobolev_seminorm}, we prove that if the density $p$ is merely upper bounded then~\eqref{eqn:graph_sobolev_seminorm_vague} holds with respect to the graph Sobolev semi-norm.
\begin{lemma}
	\label{lem:graph_sobolev_seminorm}
	Assume~\ref{asmp:kernel} and additionally that $p(x) \leq p_{\max}$ for all $x \in \Xset$. Then, for any $f \in H^1(\Xset)$, 
	\begin{equation}
	\label{eqn:graph_sobolev_seminorm}
	f^T \Lap_{n,r} f \leq \frac{p_{\max}^2 \sigma_K}{\delta} n^2 r^{d + 2} |f|_{H^1(X)}^2
	\end{equation}
	with probability at least $1 - \delta$.
\end{lemma}
A few remarks:
\begin{itemize}
	\item 
	The inequality~\eqref{eqn:graph_sobolev_seminorm} resembles~\eqref{eqn:seminorm_convergence}. However, as we pointed out previously, to derive~\eqref{eqn:seminorm_convergence} one needs $f \in C^{2}(\Xset)$ as opposed to $f \in H^1(\Xset)$, and imposing this extra degree of regularity is unsuitable for our purposes. We use a more careful analysis to prove Lemma~\ref{lem:graph_sobolev_seminorm}, but note well that~\eqref{eqn:graph_sobolev_seminorm} is only an upper bound rather than a statement about consistency. This is the price we pay for insisting that $f$ have only one derivative; fortunately an upper bound will be sufficient for our purposes in Section~\ref{sec:minimax_optimal_laplacian_smoothing}.
	\item We note also that~\eqref{eqn:graph_sobolev_seminorm}---which holds with probability $1 - \delta$---depends proportionally on $1/\delta$, as opposed to $\log(1/\delta)$. The reason: the Sobolev assumption $f \in H^1(\Xset)$ only grants us control over the second moment of $f$ and its first derivative, and so we are afforded only the relatively weak concentration implied by Markov's inequality. If we assume $f$ and its first derivative are more uniformly bounded--for instance if $f \in C^1(\Xset)$---we can use Hoeffding's inequality to get sharper upper bounds.
\end{itemize}
